---
title: "Bike Sharing Data Analysis"
output: html_notebook
---

Pre-processing

```{r}
# Clear memory
rm(list = ls())

#Loading the data in
day_data <- read.csv("day.csv")

#Forcing columns to numeric
Time = as.numeric(sub("^(\\d+).(\\d+).*", "\\1.\\2", day_data$dteday))
Date = as.Date(day_data$dteday,format = "%Y-%m-%d")
Season = as.numeric(as.character(day_data$season))
Year = as.numeric(as.character(day_data$yr))
Month = as.numeric(as.character(day_data$mnth))
Holiday = as.numeric(as.character(day_data$holiday))
Weekday = as.numeric(as.character(day_data$weekday))
WorkingDay = as.numeric(as.character(day_data$workingday))
WeatherSituation = as.numeric(as.character(day_data$weathersit))
Temperature = as.numeric(as.character(day_data$temp))
TemperatureFeel = as.numeric(as.character(day_data$atemp))
Humidity = as.numeric(as.character(day_data$hum))
WindSpeed = as.numeric(as.character(day_data$windspeed))
CountCasuals = as.numeric(as.character(day_data$casual))
CountRegistered = as.numeric(as.character(day_data$registered))
CountRentals = as.numeric(as.character(day_data$cnt))

data <- data.frame(Date,Temperature,CountRentals,Humidity, WorkingDay, Weekday, Holiday, Season, Month, WindSpeed, CountCasuals, CountRegistered, WeatherSituation)

#2011 data: training, 2012 data: validation
Training = data[(data$Date >= "2011-01-01" & data$Date <= "2011-12-31"),];
Validation = data[(data$Date >= "2012-01-01" & data$Date <= "2012-12-31"),];

#Loading necessary libraries
library(ggplot2)
library(hrbrthemes)
library(GGally)
library(glmnet)
library(ggfortify)
```

Linear Model 1

```{r}
lr_1.mls <- lm(CountRegistered ~ Temperature + WorkingDay + WindSpeed, data=Training)
summary(lr_1.mls)
mse_lr_1 <- mean(lr_1.mls$residuals^2)

lc_1.mls <- lm(CountCasuals ~ Temperature + WorkingDay + WindSpeed, data=Training)
summary(lc_1.mls)
mse_lc_1 <- mean(lc_1.mls$residuals^2)

mse_lr_1
mse_lr_1 / mean(Training$CountRegistered^2)
mse_lc_1
mse_lc_1 / mean(Training$CountCasuals^2)



ggplot() + 
  geom_point(data=data.frame(Training$CountRegistered,rstandard(lr_1.mls)), aes(x=Training$CountRegistered, y=rstandard(lr_1.mls), color = "MLS"), size = 1) +
  geom_hline(yintercept=2,color='blue') + geom_hline(yintercept=-2, color='blue') +
  scale_color_manual(name = element_blank(), labels = c("MLS"), values = c("blue")) +
  labs(y = "Standarized Residual") + ggtitle("CountRegistered Standarized Residuals Linear 1")
ggplot() + 
  geom_point(data=data.frame(Training$CountCasuals,rstandard(lc_1.mls)), aes(x=Training$CountCasuals, y=rstandard(lc_1.mls), color = "MLS"), size = 1) +
  geom_hline(yintercept=2,color='blue') + geom_hline(yintercept=-2, color='blue') +
  scale_color_manual(name = element_blank(), labels = c("MLS"), values = c("blue")) +
  labs(y = "Standarized Residual") + ggtitle("CountCasuals Standarized Residuals Linear 1")
ggplot() +
  geom_point(data=data.frame(lr_1.mls$fitted.values, rstandard(lr_1.mls)), aes(x=lr_1.mls$fitted.values, y=rstandard(lr_1.mls), color = "MLS"), size = 1) +
  geom_hline(yintercept=2,color='blue') + geom_hline(yintercept=-2, color='blue') +
  scale_color_manual(name = element_blank(), labels = c("MLS"), values = c("blue")) +
  labs(y = "Standarized Residuals") + labs(x = "Fitted values") + 
  ggtitle("CountRegistered Standarized Residuals Linear 1 (Fitted) ")
ggplot() +
  geom_point(data=data.frame(lc_1.mls$fitted.values, rstandard(lc_1.mls)), aes(x=lc_1.mls$fitted.values, y=rstandard(lc_1.mls), color = "MLS"), size = 1) +
  geom_hline(yintercept=2,color='blue') + geom_hline(yintercept=-2, color='blue') +
  scale_color_manual(name = element_blank(), labels = c("MLS"), values = c("blue")) +
  labs(y = "Standarized Residuals") + labs(x = "Fitted values") + 
  ggtitle("CountRegistered Standarized Residuals Linear 1 (Fitted) ")

p <- ggplot(data.frame(lr_1.mls$residuals), aes(sample = rstandard(lr_1.mls)))
p + stat_qq() + stat_qq_line() 

l <- ggplot(data.frame(lc_1.mls$residuals), aes(sample = rstandard(lc_1.mls)))
l + stat_qq() + stat_qq_line()


# Histograms
ggplot(data = data.frame(rstandard(lr_1.mls)), aes(x = rstandard(lr_1.mls))) + geom_histogram(bins = 30) +
  ggtitle("Histogram CountRegistered Linear 1")

ggplot(data = data.frame(rstandard(lc_1.mls)), aes(x = rstandard(lc_1.mls))) + geom_histogram(bins = 30) +
  ggtitle("Histogram CountCasuals Linear 1")
```

Linear 2

```{r}

lr_2.mls <- lm(CountRegistered ~ Temperature + WindSpeed + WorkingDay + Humidity , data=Training)
summary(lr_2.mls)
mse_lr_2 <- mean(lr_2.mls$residuals^2)

lc_2.mls <- lm(CountCasuals ~ Temperature + WindSpeed + WorkingDay + Humidity , data=Training)
summary(lc_2.mls)
mse_lc_2 <- mean(lc_2.mls$residuals^2)

pred_lr_2 <-predict(lr_2.mls, se.fit = TRUE, newdata=Validation)
vres_lr_2 <- Validation$CountRegistered - pred_lr_2$fit
pred_lc_2 <-predict(lc_2.mls, se.fit = TRUE, newdata=Validation)
vres_lc_2 <- Validation$CountCasuals - pred_lc_2$fit

mse_v_lr_2 <- mean(vres_lr_2^2)
mse_v_lc_2 <- mean(vres_lc_2^2)
rmse_lr_2 <- mse_v_lr_2 / mean(Validation$CountRegistered^2)
rmse_lc_2 <- mse_v_lc_2 / mean(Validation$CountCasuals^2)

ggplot() + 
  geom_point(data=data.frame(Training$CountRegistered,rstandard(lr_2.mls)), aes(x=Training$CountRegistered, y=rstandard(lr_2.mls), color = "MLS"), size = 1) +
  geom_hline(yintercept=2,color='blue') + geom_hline(yintercept=-2, color='blue') +
  scale_color_manual(name = element_blank(), labels = c("MLS"), values = c("blue")) +
  labs(y = "Standarized Residual") + ggtitle("CountRegistered Standarized Residuals Linear 1")
ggplot() + 
  geom_point(data=data.frame(Training$CountCasuals,rstandard(lc_2.mls)), aes(x=Training$CountCasuals, y=rstandard(lc_2.mls), color = "MLS"), size = 1) +
  geom_hline(yintercept=2,color='blue') + geom_hline(yintercept=-2, color='blue') +
  scale_color_manual(name = element_blank(), labels = c("MLS"), values = c("blue")) +
  labs(y = "Standarized Residual") + ggtitle("CountCasuals Standarized Residuals Linear 1")
ggplot() +
  geom_point(data=data.frame(lr_2.mls$fitted.values, rstandard(lr_2.mls)), aes(x=lr_2.mls$fitted.values, y=rstandard(lr_2.mls), color = "MLS"), size = 1) +
  geom_hline(yintercept=2,color='blue') + geom_hline(yintercept=-2, color='blue') +
  scale_color_manual(name = element_blank(), labels = c("MLS"), values = c("blue")) +
  labs(y = "Standarized Residuals") + labs(x = "Fitted values") + 
  ggtitle("CountRegistered Standarized Residuals Linear 1 (Fitted) ")
ggplot() +
  geom_point(data=data.frame(lc_2.mls$fitted.values, rstandard(lc_2.mls)), aes(x=lc_2.mls$fitted.values, y=rstandard(lc_2.mls), color = "MLS"), size = 1) +
  geom_hline(yintercept=2,color='blue') + geom_hline(yintercept=-2, color='blue') +
  scale_color_manual(name = element_blank(), labels = c("MLS"), values = c("blue")) +
  labs(y = "Standarized Residuals") + labs(x = "Fitted values") + 
  ggtitle("CountRegistered Standarized Residuals Linear 1 (Fitted) ")
p <- ggplot(data.frame(lr_2.mls$residuals), aes(sample = rstandard(lr_2.mls)))
p + stat_qq() + stat_qq_line() 

l <- ggplot(data.frame(lc_2.mls$residuals), aes(sample = rstandard(lc_2.mls)))
l + stat_qq() + stat_qq_line()


# Histograms
ggplot(data = data.frame(rstandard(lr_2.mls)), aes(x = rstandard(lr_2.mls))) + geom_histogram(bins = 30) +
  ggtitle("Histogram Registered MLS Plot")

ggplot(data = data.frame(rstandard(lc_2.mls)), aes(x = rstandard(lc_2.mls))) + geom_histogram(bins = 30) +
  ggtitle("Histogram Casual MLS Plot")

```

Linear 3

```{r}

lr_3.mls <- lm(CountRegistered ~ Temperature + Weekday + WeatherSituation, data=Training)
summary(lr_3.mls)
mse_lr_3 <- mean(lr_3.mls$residuals^2)

lc_3.mls <- lm(CountCasuals ~ Temperature + Weekday + WeatherSituation, data=Training)
summary(lc_3.mls)
mse_lc_3 <- mean(lc_3.mls$residuals^2)

pred_lr_3 <-predict(lr_3.mls, se.fit = TRUE, newdata=Validation)
vres_lr_3 <- Validation$CountRegistered - pred_lr_3$fit
pred_lc_3 <-predict(lc_3.mls, se.fit = TRUE, newdata=Validation)
vres_lc_3 <- Validation$CountCasuals - pred_lc_3$fit

mse_v_lr_3 <- mean(vres_lr_3^2)
mse_v_lc_3 <- mean(vres_lc_3^2)
rmse_lr_3 <- mse_v_lr_3 / mean(Validation$CountRegistered^2)
rmse_lc_3 <- mse_v_lc_3 / mean(Validation$CountCasuals^2)

ggplot() + 
  geom_point(data=data.frame(Training$CountRegistered,rstandard(lr_3.mls)), aes(x=Training$CountRegistered, y=rstandard(lr_3.mls), color = "MLS"), size = 1) +
  geom_hline(yintercept=2,color='blue') + geom_hline(yintercept=-2, color='blue') +
  scale_color_manual(name = element_blank(), labels = c("MLS"), values = c("blue")) +
  labs(y = "Standarized Residual") + ggtitle("CountRegistered Standarized Residuals Linear 1")
ggplot() + 
  geom_point(data=data.frame(Training$CountCasuals,rstandard(lc_3.mls)), aes(x=Training$CountCasuals, y=rstandard(lc_3.mls), color = "MLS"), size = 1) +
  geom_hline(yintercept=2,color='blue') + geom_hline(yintercept=-2, color='blue') +
  scale_color_manual(name = element_blank(), labels = c("MLS"), values = c("blue")) +
  labs(y = "Standarized Residual") + ggtitle("CountCasuals Standarized Residuals Linear 1")
ggplot() +
  geom_point(data=data.frame(lr_3.mls$fitted.values, rstandard(lr_3.mls)), aes(x=lr_3.mls$fitted.values, y=rstandard(lr_3.mls), color = "MLS"), size = 1) +
  geom_hline(yintercept=2,color='blue') + geom_hline(yintercept=-2, color='blue') +
  scale_color_manual(name = element_blank(), labels = c("MLS"), values = c("blue")) +
  labs(y = "Standarized Residuals") + labs(x = "Fitted values") + 
  ggtitle("CountRegistered Standarized Residuals Linear 1 (Fitted) ")
ggplot() +
  geom_point(data=data.frame(lc_3.mls$fitted.values, rstandard(lc_3.mls)), aes(x=lc_3.mls$fitted.values, y=rstandard(lc_3.mls), color = "MLS"), size = 1) +
  geom_hline(yintercept=2,color='blue') + geom_hline(yintercept=-2, color='blue') +
  scale_color_manual(name = element_blank(), labels = c("MLS"), values = c("blue")) +
  labs(y = "Standarized Residuals") + labs(x = "Fitted values") + 
  ggtitle("CountRegistered Standarized Residuals Linear 1 (Fitted) ")

p <- ggplot(data.frame(lr_3.mls$residuals), aes(sample = rstandard(lr_3.mls)))
p + stat_qq() + stat_qq_line() 

l <- ggplot(data.frame(lc_3.mls$residuals), aes(sample = rstandard(lc_3.mls)))
l + stat_qq() + stat_qq_line()


# Histograms
ggplot(data = data.frame(rstandard(lr_3.mls)), aes(x = rstandard(lr_3.mls))) + geom_histogram(bins = 30) +
  ggtitle("Histogram Registered MLS Plot")

ggplot(data = data.frame(rstandard(lc_3.mls)), aes(x = rstandard(lc_3.mls))) + geom_histogram(bins = 30) +
  ggtitle("Histogram Casual MLS Plot")

```

Quadratic 1

```{r}

qr_1.mls <- lm(CountRegistered~Temperature + I((Temperature)^2)+WindSpeed+WorkingDay, data=Training)
summary(qr_1.mls)
mse_qr_1 <- mean(qr_1.mls$residuals^2)

qc_1.mls <- lm(CountCasuals~Temperature + I((Temperature)^2)+WindSpeed+WorkingDay, data=Training)
summary(qc_1.mls)
mse_qc_1 <- mean(qc_1.mls$residuals^2)

mse_qr_1
mse_qr_1 / mean(Training$CountRegistered^2)
mse_qc_1
mse_qc_1 / mean(Training$CountCasuals^2)

ggplot() + 
  geom_point(data=data.frame(Training$CountRegistered,rstandard(qr_1.mls)), aes(x=Training$CountRegistered, y=rstandard(qr_1.mls), color = "MLS"), size = 1) +
  geom_hline(yintercept=2,color='blue') + geom_hline(yintercept=-2, color='blue') +
  scale_color_manual(name = element_blank(), labels = c("MLS"), values = c("blue")) +
  labs(y = "Standarized Residual") + ggtitle("CountRegistered Standarized Residuals Linear 1")
ggplot() + 
  geom_point(data=data.frame(Training$CountCasuals,rstandard(qc_1.mls)), aes(x=Training$CountCasuals, y=rstandard(qc_1.mls), color = "MLS"), size = 1) +
  geom_hline(yintercept=2,color='blue') + geom_hline(yintercept=-2, color='blue') +
  scale_color_manual(name = element_blank(), labels = c("MLS"), values = c("blue")) +
  labs(y = "Standarized Residual") + ggtitle("CountCasuals Standarized Residuals Linear 1")
ggplot() +
  geom_point(data=data.frame(qr_1.mls$fitted.values, rstandard(qr_1.mls)), aes(x=qr_1.mls$fitted.values, y=rstandard(qr_1.mls), color = "MLS"), size = 1) +
  geom_hline(yintercept=2,color='blue') + geom_hline(yintercept=-2, color='blue') +
  scale_color_manual(name = element_blank(), labels = c("MLS"), values = c("blue")) +
  labs(y = "Standarized Residuals") + labs(x = "Fitted values") + 
  ggtitle("CountRegistered Standarized Residuals Linear 1 (Fitted) ")
ggplot() +
  geom_point(data=data.frame(qc_1.mls$fitted.values, rstandard(qc_1.mls)), aes(x=qc_1.mls$fitted.values, y=rstandard(qc_1.mls), color = "MLS"), size = 1) +
  geom_hline(yintercept=2,color='blue') + geom_hline(yintercept=-2, color='blue') +
  scale_color_manual(name = element_blank(), labels = c("MLS"), values = c("blue")) +
  labs(y = "Standarized Residuals") + labs(x = "Fitted values") + 
  ggtitle("CountRegistered Standarized Residuals Linear 1 (Fitted) ")

p <- ggplot(data.frame(qr_1.mls$residuals), aes(sample = rstandard(qr_1.mls)))
p + stat_qq() + stat_qq_line() 

l <- ggplot(data.frame(qc_1.mls$residuals), aes(sample = rstandard(qc_1.mls)))
l + stat_qq() + stat_qq_line()


# Histograms
ggplot(data = data.frame(rstandard(qr_1.mls)), aes(x = rstandard(qr_1.mls))) + geom_histogram(bins = 30) +
  ggtitle("Histogram Registered MLS Plot")

ggplot(data = data.frame(rstandard(qc_1.mls)), aes(x = rstandard(qc_1.mls))) + geom_histogram(bins = 30) +
  ggtitle("Histogram Casual MLS Plot")
```

Quadratic 2

```{r}

qr_2.mls <- lm(CountRegistered ~ Temperature + I((Temperature)^2) + WindSpeed + I((WindSpeed)^2) + WorkingDay + Humidity + I((Humidity)^2), data=Training)
summary(qr_2.mls)
mse_qr_2 <- mean(qr_2.mls$residuals^2)

qc_2.mls <- lm(CountCasuals ~ Month + I((Month)^2) + Temperature + I((Temperature)^2) + WindSpeed + I((WindSpeed)^2) + WorkingDay + Humidity + I((Humidity)^2), data=Training)
summary(qc_2.mls)
mse_qc_2 <- mean(qc_2.mls$residuals^2)

pred_qr_2 <-predict(qr_2.mls, se.fit = TRUE, newdata=Validation)
vres_qr_2 <- Validation$CountRegistered - pred_qr_2$fit
pred_qc_2 <-predict(qc_2.mls, se.fit = TRUE, newdata=Validation)
vres_qc_2 <- Validation$CountCasuals - pred_qc_2$fit

mse_v_qr_2 <- mean(vres_qr_2^2)
mse_v_qc_2 <- mean(vres_qc_2^2)
rmse_qr_2 <- mse_v_qr_2 / mean(Validation$CountRegistered^2)
rmse_qc_2 <- mse_v_qc_2 / mean(Validation$CountCasuals^2)

ggplot() + 
  geom_point(data=data.frame(Training$CountRegistered,rstandard(qr_2.mls)), aes(x=Training$CountRegistered, y=rstandard(qr_2.mls), color = "MLS"), size = 1) +
  geom_hline(yintercept=2,color='blue') + geom_hline(yintercept=-2, color='blue') +
  scale_color_manual(name = element_blank(), labels = c("MLS"), values = c("blue")) +
  labs(y = "Standarized Residual") + ggtitle("CountRegistered Standarized Residuals Linear 1")
ggplot() + 
  geom_point(data=data.frame(Training$CountCasuals,rstandard(qc_2.mls)), aes(x=Training$CountCasuals, y=rstandard(qc_2.mls), color = "MLS"), size = 1) +
  geom_hline(yintercept=2,color='blue') + geom_hline(yintercept=-2, color='blue') +
  scale_color_manual(name = element_blank(), labels = c("MLS"), values = c("blue")) +
  labs(y = "Standarized Residual") + ggtitle("CountCasuals Standarized Residuals Linear 1")
ggplot() +
  geom_point(data=data.frame(qr_2.mls$fitted.values, rstandard(qr_2.mls)), aes(x=qr_2.mls$fitted.values, y=rstandard(qr_2.mls), color = "MLS"), size = 1) +
  geom_hline(yintercept=2,color='blue') + geom_hline(yintercept=-2, color='blue') +
  scale_color_manual(name = element_blank(), labels = c("MLS"), values = c("blue")) +
  labs(y = "Standarized Residuals") + labs(x = "Fitted values") + 
  ggtitle("CountRegistered Standarized Residuals Linear 1 (Fitted) ")
ggplot() +
  geom_point(data=data.frame(qc_2.mls$fitted.values, rstandard(qc_2.mls)), aes(x=qc_2.mls$fitted.values, y=rstandard(qc_2.mls), color = "MLS"), size = 1) +
  geom_hline(yintercept=2,color='blue') + geom_hline(yintercept=-2, color='blue') +
  scale_color_manual(name = element_blank(), labels = c("MLS"), values = c("blue")) +
  labs(y = "Standarized Residuals") + labs(x = "Fitted values") + 
  ggtitle("CountRegistered Standarized Residuals Linear 1 (Fitted) ")

p <- ggplot(data.frame(qr_2.mls$residuals), aes(sample = rstandard(qr_2.mls)))
p + stat_qq() + stat_qq_line() 

l <- ggplot(data.frame(qc_2.mls$residuals), aes(sample = rstandard(qc_2.mls)))
l + stat_qq() + stat_qq_line()


# Histograms
ggplot(data = data.frame(rstandard(qr_2.mls)), aes(x = rstandard(qr_2.mls))) + geom_histogram(bins = 30) +
  ggtitle("Histogram Registered MLS Plot")

ggplot(data = data.frame(rstandard(qc_2.mls)), aes(x = rstandard(qc_2.mls))) + geom_histogram(bins = 30) +
  ggtitle("Histogram Casual MLS Plot")
```

Quadratic 3

```{r}

qr_3.mls <- lm(CountRegistered ~ Temperature + I((Temperature)^2) + Weekday + WeatherSituation, data=Training)
summary(qr_2.mls)
mse_qr_3 <- mean(qr_2.mls$residuals^2)

qc_3.mls <- lm(CountCasuals ~ Temperature + I((Temperature)^2) + Weekday + WeatherSituation, data=Training)
summary(qc_3.mls)
mse_qc_3 <- mean(qc_3.mls$residuals^2)

pred_qr_3 <-predict(qr_3.mls, se.fit = TRUE, newdata=Validation)
vres_qr_3 <- Validation$CountRegistered - pred_qr_3$fit
pred_qc_3 <-predict(qc_3.mls, se.fit = TRUE, newdata=Validation)
vres_qc_3 <- Validation$CountCasuals - pred_qc_3$fit

mse_v_qr_3 <- mean(vres_qr_3^2)
mse_v_qc_3 <- mean(vres_qc_3^2)
rmse_qr_3 <- mse_v_qr_3 / mean(Validation$CountRegistered^2)
rmse_qc_3 <- mse_v_qc_3 / mean(Validation$CountCasuals^2)

ggplot() + 
  geom_point(data=data.frame(Training$CountRegistered,rstandard(qr_3.mls)), aes(x=Training$CountRegistered, y=rstandard(qr_3.mls), color = "MLS"), size = 1) +
  geom_hline(yintercept=2,color='blue') + geom_hline(yintercept=-2, color='blue') +
  scale_color_manual(name = element_blank(), labels = c("MLS"), values = c("blue")) +
  labs(y = "Standarized Residual") + ggtitle("CountRegistered Standarized Residuals Linear 1")
ggplot() + 
  geom_point(data=data.frame(Training$CountCasuals,rstandard(qc_3.mls)), aes(x=Training$CountCasuals, y=rstandard(qc_3.mls), color = "MLS"), size = 1) +
  geom_hline(yintercept=2,color='blue') + geom_hline(yintercept=-2, color='blue') +
  scale_color_manual(name = element_blank(), labels = c("MLS"), values = c("blue")) +
  labs(y = "Standarized Residual") + ggtitle("CountCasuals Standarized Residuals Linear 1")
ggplot() +
  geom_point(data=data.frame(qr_3.mls$fitted.values, rstandard(qr_3.mls)), aes(x=qr_3.mls$fitted.values, y=rstandard(qr_3.mls), color = "MLS"), size = 1) +
  geom_hline(yintercept=2,color='blue') + geom_hline(yintercept=-2, color='blue') +
  scale_color_manual(name = element_blank(), labels = c("MLS"), values = c("blue")) +
  labs(y = "Standarized Residuals") + labs(x = "Fitted values") + 
  ggtitle("CountRegistered Standarized Residuals Linear 1 (Fitted) ")
ggplot() +
  geom_point(data=data.frame(qc_3.mls$fitted.values, rstandard(qc_3.mls)), aes(x=qc_3.mls$fitted.values, y=rstandard(qc_3.mls), color = "MLS"), size = 1) +
  geom_hline(yintercept=2,color='blue') + geom_hline(yintercept=-2, color='blue') +
  scale_color_manual(name = element_blank(), labels = c("MLS"), values = c("blue")) +
  labs(y = "Standarized Residuals") + labs(x = "Fitted values") + 
  ggtitle("CountRegistered Standarized Residuals Linear 1 (Fitted) ")

p <- ggplot(data.frame(qr_3.mls$residuals), aes(sample = rstandard(qr_3.mls)))
p + stat_qq() + stat_qq_line() 

l <- ggplot(data.frame(qc_3.mls$residuals), aes(sample = rstandard(qc_3.mls)))
l + stat_qq() + stat_qq_line()


# Histograms
ggplot(data = data.frame(rstandard(qr_3.mls)), aes(x = rstandard(qr_3.mls))) + geom_histogram(bins = 30) +
  ggtitle("Histogram Registered MLS Plot")

ggplot(data = data.frame(rstandard(qc_3.mls)), aes(x = rstandard(qc_3.mls))) + geom_histogram(bins = 30) +
  ggtitle("Histogram Casual MLS Plot")

```
MSE and rMSE comparisons

```{r}

"linear 1"
mse_v_lr_1
mse_lc_1
rmse_lr_1
rmse_lc_1
"linear 2"
mse_v_lr_2
mse_lc_2
rmse_lr_2
rmse_lc_2
"linear 3"
mse_v_lr_3
mse_lc_3
rmse_lr_3
rmse_lc_3
"quadratic 1"
mse_v_qr_1
mse_v_qc_1
rmse_qr_1
rmse_qc_1
"quadratic 2"
mse_v_qr_2
mse_v_qc_2
rmse_qr_2
rmse_qc_2
"quadratic 3"
mse_v_qr_3
mse_v_qc_3
rmse_qr_3
rmse_qc_3

#95% confidence intervals
round(confint(r_model.mls,level=0.95),7)
round(confint(c_model.mls,level=0.95),7)


# Covariance([bo,b1])
vcov(r_model.mls)
vcov(c_model.mls)
```


Validation


```{r}

#Quadratic 1

pred_qr_1 <-predict(qr_1.mls, se.fit = TRUE, newdata=Validation)
vres_qr_1 <- Validation$CountRegistered - pred_qr_1$fit
pred_qc_1<- predict(qc_1.mls, se.fit = TRUE, newdata=Validation)
vres_qc_1 <- Validation$CountCasuals - pred_qc_1$fit

mse_v_qr_1 <- mean(vres_qr_1^2)
mse_v_qc_1 <- mean(vres_qc_1^2)
rmse_qr_1 <- mse_v_qr_1 / mean(Validation$CountRegistered^2)
rmse_qc_1 <- mse_v_qc_1 / mean(Validation$CountCasuals^2)

"qr_1"
mse_v_qr_1
mse_qr_1
rmse_qr_1
"qc_1"
mse_v_qc_1
mse_qc_1
rmse_qc_1

#Plotting actual vs predicted for CountRegistered

# Create data frame with validation observation and prediction
test_qr_1 = data.frame(Validation$CountRegistered,pred_qr_1$fit, 1:length(pred_qr_1$fit));
colnames(test_qr_1)[1] = "CountRegistered"
colnames(test_qr_1)[2] = "Prediction"
colnames(test_qr_1)[3] = "Index"

ggplot(data = test_qr_1, aes(x = CountRegistered, y = Prediction)) + geom_point() + 
  geom_abline(intercept = 0, slope = 1) +
  ggtitle("Validation Quadratic CountRegistered vs Prediction")

# Further comparisons
ggplot(data = test_qr_1, aes(x = Index)) +
  geom_line(aes(y = CountRegistered, color = "CountRegistered")) + 
  geom_line(aes(y = Prediction, color="Prediction"), linetype="twodash") +  
  scale_color_manual(name = element_blank(), labels = c("CountRegistered","Prediction"),
                     values = c("darkred", "steelblue")) + labs(y = "") + 
  ggtitle("Quadratic CountRegistered Validation")

# Hard to see, let's zoom in
test2_qr_1 = test_qr_1[100:150,]

ggplot(data = test2_qr_1, aes(x = Index)) +
  geom_line(aes(y = CountRegistered, color = "CountRegistered")) + 
  geom_line(aes(y = Prediction, color="Prediction"), linetype="twodash") +  
  scale_color_manual(name = element_blank(), labels = c("CountRegistered","Prediction"),
                     values = c("darkred", "steelblue")) + labs(y = "") +
  ggtitle("Quadratic CountRegistered Validation")


#Plotting actual vs predicted for CountCasuals

# Create data frame with validation observation and prediction
test_qc_1 = data.frame(Validation$CountCasuals,pred_qc_1$fit, 1:length(pred_qc_1$fit));
colnames(test_qc_1)[1] = "CountCasuals"
colnames(test_qc_1)[2] = "Prediction"
colnames(test_qc_1)[3] = "Index"

ggplot(data = test_qc_1, aes(x = CountCasuals, y = Prediction)) + geom_point() + 
  geom_abline(intercept = 0, slope = 1) +
  ggtitle("Validation Quadratic CountCasuals vs Prediction")

# Further comparisons
ggplot(data = test_qc_1, aes(x = Index)) +
  geom_line(aes(y = CountCasuals, color = "CountCasuals")) + 
  geom_line(aes(y = Prediction, color="Prediction"), linetype="twodash") +  
  scale_color_manual(name = element_blank(), labels = c("CountCasuals","Prediction"),
                     values = c("darkred", "steelblue")) + labs(y = "") + 
  ggtitle("Quadratic Validation CountCasuals")

# Hard to see, let's zoom in
test2_qc_1 = test_qc_1[100:150,]

ggplot(data = test2_qc_1, aes(x = Index)) +
  geom_line(aes(y = CountCasuals, color = "CountCasuals")) + 
  geom_line(aes(y = Prediction, color="Prediction"), linetype="twodash") +  
  scale_color_manual(name = element_blank(), labels = c("CountCasuals","Prediction"),
                     values = c("darkred", "steelblue")) + labs(y = "") +
  ggtitle("Quadratic Validation CountCasuals")


#Linear 1

pred_lr_1 <-predict(lr_1.mls, se.fit = TRUE, newdata=Validation)
vres_lr_1 <- Validation$CountRegistered - pred_lr_1$fit
pred_lc_1 <-predict(lc_1.mls, se.fit = TRUE, newdata=Validation)
vres_lc_1 <- Validation$CountCasuals - pred_lc_1$fit

mse_v_lr_1 <- mean(vres_lr_1^2)
mse_v_lc_1 <- mean(vres_lc_1^2)
rmse_lr_1 <- mse_v_lr_1 / mean(Validation$CountRegistered^2)
rmse_lc_1 <- mse_v_lc_1 / mean(Validation$CountCasuals^2)

"lr_1"
mse_v_lr_1
mse_lr_1
rmse_lr_1
"lc_1"
mse_v_lc_1
mse_lc_1
rmse_lc_1

test_lr_1 = data.frame(Validation$CountRegistered,pred_lr_1$fit, 1:length(pred_lr_1$fit));
colnames(test_lr_1)[1] = "CountRegistered"
colnames(test_lr_1)[2] = "Prediction"
colnames(test_lr_1)[3] = "Index"

ggplot(data = test_lr_1, aes(x = CountRegistered, y = Prediction)) + geom_point() + 
  geom_abline(intercept = 0, slope = 1) +
  ggtitle("Linear Validation CountRegistered vs Prediction")

# Further comparisons
ggplot(data = test_lr_1, aes(x = Index)) +
  geom_line(aes(y = CountRegistered, color = "CountRegistered")) + 
  geom_line(aes(y = Prediction, color="Prediction"), linetype="twodash") +  
  scale_color_manual(name = element_blank(), labels = c("CountRegistered","Prediction"),
                     values = c("darkred", "steelblue")) + labs(y = "") + 
  ggtitle("Linear Validation CountRegistered")

# Hard to see, let's zoom in
test2_lr_1 = test_lr_1[100:150,]

ggplot(data = test2_lr_1, aes(x = Index)) +
  geom_line(aes(y = CountRegistered, color = "CountRegistered")) + 
  geom_line(aes(y = Prediction, color="Prediction"), linetype="twodash") +  
  scale_color_manual(name = element_blank(), labels = c("CountRegistered","Prediction"),
                     values = c("darkred", "steelblue")) + labs(y = "") +
  ggtitle("Linear Validation CountRegistered")


#Plotting actual vs predicted for CountCasuals

# Create data frame with validation observation and prediction
test_lc_1 = data.frame(Validation$CountCasuals,pred_lc_1$fit, 1:length(pred_lc_1$fit));
colnames(test_lc_1)[1] = "CountCasuals"
colnames(test_lc_1)[2] = "Prediction"
colnames(test_lc_1)[3] = "Index"

ggplot(data = test_lc_1, aes(x = CountCasuals, y = Prediction)) + geom_point() + 
  geom_abline(intercept = 0, slope = 1) +
  ggtitle("Validation CountCasuals vs Prediction")

# Further comparisons
ggplot(data = test_lc_1, aes(x = Index)) +
  geom_line(aes(y = CountCasuals, color = "CountCasuals")) + 
  geom_line(aes(y = Prediction, color="Prediction"), linetype="twodash") +  
  scale_color_manual(name = element_blank(), labels = c("CountCasuals","Prediction"),
                     values = c("darkred", "steelblue")) + labs(y = "") + 
  ggtitle("Linear Validation CountCasuals")

# Hard to see, let's zoom in
test2_lc_1 = test_lc_1[100:150,]

ggplot(data = test2_lc_1, aes(x = Index)) +
  geom_line(aes(y = CountCasuals, color = "CountCasuals")) + 
  geom_line(aes(y = Prediction, color="Prediction"), linetype="twodash") +  
  scale_color_manual(name = element_blank(), labels = c("CountCasuals","Prediction"),
                     values = c("darkred", "steelblue")) + labs(y = "") +
  ggtitle("Linear Validation CountCasuals")


```

Takeaways from Validation:

We need to train on multi-year data to accurately account for growth of the bike rental market not related to weather/situational conditions. 
If we acquired future data, it would make sense to train and cross validate on multi-year data, because it's possible that didn't fit the 2011 data as well would have fit the 2012 data better (possibly run some analysis if Julio says that's ok)





We don't have information on the number of registered riders (just how much they rented) so likely that customer base just grew (because casual riders remained relatively the same)



